A Dashboard for the Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-09 discussion.tex
Conclusions updated
lay outing...

\section{Discussion}

@@

    This technique assumes that the web application code which is the target of the monitoring is deployed using \git in the following way:

    \begin{enumerate}
      \item The deployment engineer pulls the latest version of the code from the integration server; this will result in a new commit being pointed at by the \code{HEAD} pointer. %than previously
      \item The deployment engineer restarts the new version of the service. At this point, the \tool detects that a new \code{HEAD} is present in the local code base and consequently starts associating all the new data points with this new commit\footnote{The \tool detects the current version of the analyzed system the first time it is attached to the application object, and thus, assumes that the Flask application is restarted when a new version is deployed. This is in tune with the current version of Flask, but if the web server will support dynamic updates in the future, this might have to be taken into account}.
    \end{enumerate}

      The advantage of this approach is the need for minimal configuration effort, as discussed in the presentation of the \tool. The disadvantage is that it will consider on equal ground the smallest of commits, even one that modifies a comment, and the shortest lived of commits, e.g.~a commit which was active only for a half an hour before a new version with a bug fix was deployed, with major and minor releases of the software. %as a distinct way of grouping the data points. 
    A mechanism to control which versions are important for monitoring purposes is therefore required to be added to the \tool.

>>>>

\section{Discussion}
\label{sec:discussion}

@@

      The advantage of the presented approach to evolution monitoring based on observing the \code{.git} folder is the need for minimal configuration effort, as discussed in the presentation of the \tool. The disadvantage is that it will consider on equal ground the smallest of commits, even one that modifies a comment, and the shortest lived of commits, e.g.~a commit which was active only for a half an hour before a new version with a bug fix was deployed, with major and minor releases of the software. %as a distinct way of grouping the data points. 
    A mechanism to control which versions are important for monitoring purposes is therefore required to be added to the \tool.

### 2018-04-09 evolution.tex
...
anonymization
lay outing...

\begin{lstlisting}[style=custompython]

# LOC #3: provide the dashboard with 
# information on where to find 
# git information 
dashboard.config.git = 'path/to/.git'

@@

@@

@@

    \begin{figure}
      \centering
      \includegraphics[width=0.9\linewidth]{utilization-evolution}
      \caption{The Evolution of All the API Endpoints Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\linewidth]{translation_endpoint_evolution_}
      \caption{The Performance Evolution of the \epTranslations endpoint}
      \label{fig:tee}
    \end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.77\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis) for a given endpoint}
  \label{fig:tuv}
\end{figure}

>>>>

\begin{lstlisting}[style=custompython]
# LOC #3: provide the dashboard with 
# information on where to find 
# git information 
dashboard.config.git = 'path/to/.git'

@@

    This technique assumes that the web application code which is the target of the monitoring is deployed using \git by pulling the latest version of the code from the integration server; this will result in a new commit being pointed at by the \code{HEAD} pointer. Thus, when the deployment engineer restarts the new version of the service, the \tool detects that a new \code{HEAD} is present in the local code base and consequently starts associating all the new data points with this new commit. The \tool detects the current version of the analyzed system the first time it is attached to the application object, and thus, assumes that the Flask application is restarted when a new version is deployed. This would not work in the presence of dynamic updates.

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\linewidth]{utilization-evolution}
      \caption{Evolving Endpoint Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

@@

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.97\linewidth]{translation_endpoint_evolution_}
      \caption{Endpoint performance across three versions}
      \label{fig:tee}
    \end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis) for a given endpoint}
  \label{fig:tuv}
\end{figure}

### 2018-04-09 grouping.tex
lay outing...

\begin{lstlisting}[style=custompython]  
# LOC #2: configure the dashboard
# to group requests by the user id
dashboard.config.group_by = 'User',
  lambda: Session.find(flask.request).user.id

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

>>>>

\begin{lstlisting}[style=custompython]  
# LOC #2: group requests by user id
dashboard.config.group_by = 'User',
  lambda: Session.find(flask.request).user.id

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.66\linewidth]{time_per_user}
  \caption{And endpoint with a high variability across users}
  \label{fig:tpu}
\end{figure}

### 2018-04-09 testing.tex
a nice table with the benchmark numbers is in
lay outing...

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.65\columnwidth]{travis_builds-bw}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

@@

  \subsection*{Preemptive Monitoring}
  The concept of {\em preemptive monitoring} of the application performance by means of instrumenting integration unit testing as the synthetic load is similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of the CI framework to create an emulated ``live'' environment for integration testing purposes, and use unit testing as the dedicated test input in order to measure performance.

  While this integration testing environment is different from the production one, and the load used is purely synthetic, it can serve as an early performance indicator for the developer.  \Fref{fig:response_times_preemptive} shows the actual endpoint response times across five versions of the system (above) and the corresponding testing times for the synthetic load in the same five versions (below). The green line joining the medians hints to a correlation between the two sets of measurements.

@@

        \advance\leftskip-0.2cm
        \includegraphics[width=0.5\columnwidth]{travis_builds_no_outliers_trend}
        \caption{The reported response time (in ms) per deployed version in the observation period: 
        in the actual deployment (above) and using integration with Travis (below)}

@@

  Computing Pearson correlation ($r(3)=.93, p=.02$) between the median values of the two datasets (cf. Table \ref{tab:correlations}) shows that there is indeed a correlation between the two. Further investigation on the characteristics of this observation is currently ongoing and it will be elaborated further in future work.

>>>>

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.85\columnwidth]{travis_builds-bw}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

@@

  \subsection*{Preemptive Monitoring}
  The concept of {\em preemptive monitoring} of the application performance by means of instrumenting integration unit testing as the synthetic load is similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of the CI framework to create an emulated ``live'' environment for integration testing purposes, and use unit testing as the dedicated test input. % in order to measure performance.

  While this integration testing environment is different from the production one, and the load used is purely synthetic, it can serve as an early performance indicator for the developer.  \Fref{fig:response_times_preemptive} shows the actual endpoint response times across five versions of the system (above) and the corresponding testing times for the synthetic load in the same five versions (below).

@@

        \advance\leftskip-0.2cm
        \includegraphics[width=0.5\columnwidth]{travis_builds_no_outliers_trend}
        \caption{The reported response time (in ms) per deployed version in the observation period: 
        in the actual deployment (above) and using integration with Travis (below). 
        Green line hints to a correlation between the two sets of measurements. }

@@

  Computing Pearson correlation ($r(3)=.93, p=.02$) between the median values of the two datasets (cf. Table \ref{tab:correlations}) shows a correlation between the two. Further investigation on the characteristics of this observation is currently ongoing and it will be elaborated further in future work.
